### 2023-09-27 submission/supplemental_text.Rmd

---
bibliography: references.bib
output:
  pdf_document:
    keep_tex: true
csl: asm.csl
geometry: margin=1.0in
header-includes:
 - \usepackage{setspace}
 - \doublespacing
 - \usepackage{helvet} % Helvetica font
 - \renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
 - \usepackage[T1]{fontenc}
 - \usepackage[shortcuts]{extdash}
 - \usepackage[document]{ragged2e}
---

## Supplementary text

Quotes from selected papers that cited WNWN and conflated "rarefying"/"rarefy" and "rarefaction". Many of these papers also attributed "rarefaction" to WNWN. Rarefying, rarefy, and rarefaction are bolded to make the words easier to identify. All references have been updated to reflect the numbering in the current study. Citations to WNWN are indicated by [@McMurdie2014]. Quotes are ordered chronologically.

* "Additionally, **rarefaction** has recently been shown to introduce errors in analyses, and alternatives to rarefaction have been proposed [@McMurdie2014]" - Quoted from [@Goodrich2014]. Implies that rarefaction and WNWN's rarefying are the same thing.

* "**Rarefaction** is analytically problematic and poses multiple statistical problems: (i) omission of available valid data, (ii) the estimation of overdispersion is more difficult due to data loss, (iii) loss of power (type II error), (iv) dependence on an arbitrary threshold and (v) additional uncertainty due to the randomness in **rarefaction** [@McMurdie2014]." - Quoted from [@Blint2016]. Implies that rarefaction and WNWN's rarefying are the same thing.

* "McMurdie and Holmes [@McMurdie2014] penalized the **rarefying** technique for dropping the lowest fifteenth percentile of sample library sizes in their simulations by counting the dropped samples as "incorrectly clustered." Because the 15th percentile was used to set **rarefaction** depth, this capped clustering accuracy at 85%." - Quoted from [@Weiss2017]. An example of a paper from the QIIME development community using rarefaction and rarefying interchangeably.

* "**Rarefaction** has a limited ability in this regard since the total sum constraint still exists after **rarefaction**. In addition, it suffers from a great power loss due to the discard of a large number of reads [@McMurdie2014]." - Quoted from [@Chen2018]. Implies that rarefaction and WNWN's rarefying are the same thing.

* "Another popular normalization approach is **rarefaction**, which consists on subsampling the same number of reads for each sample so that all samples have the same number of total counts. **Rarefaction** is not recommended because it entails the loss of important information [@McMurdie2014]." - Quoted from [@Calle2019]. Implies that rarefaction and WNWN's rarefying are the same thing.

* "Unfortunately, **rarefaction** is neither justifiable nor necessary, a view framed statistically by McMurdie and Holmes [@McMurdie2014] in the context of comparison of relative abundances." - Quoted from [@Willis2019]. Implies that rarefaction and WNWN's rarefying are the same thing.

* "Some studies perform **rarefaction** to adjust for differences in library size due to unexhaustive metagenomic sampling. Although several pipelines provide this functionality, it has been found inadmissible for metagenomics microbiome studies as it discards many reads leading to decreased sensitivity in differential abundance testing [@McMurdie2014] and biased estimates for alpha diversity [@Willis2019]." - Quoted from [@Eetemadi2020]. Implies that rarefaction and WNWN's rarefying are the same thing.

* "Another widespread practice is to **rarefy** the count data to force the samples to have the same number of total sequence reads [@Hughes2005], at the expense of discarding vast amounts of information [@McMurdie2014]." - Quoted from [@Leite2020]. Hughes and Hellmann [@Hughes2005] describe traditional rarefaction with multiple subsamplings.

* "Note that the terms **rarefying** and **rarefaction** are used interchangeably in microbiome literature [@McMurdie2014]. **Rarefying** was first recommended for microbiome data to deal with rare taxa [@Lozupone2010], which impact some measures of alpha and beta diversities [@McMurdie2014]." - Quoted from [@Lin2020]. As the quote indicates they use rarefying and rarefaction interchangeably although the cited WNWN indicates that they are not the same. In addition, the first review article to advocate the use of rarefaction was Hughes et al. [@Hughes2001] and examples of its use can be found earlier [@Dunbar1999; @Moyer1998];

* "Ways to tackle this include total library size normalization and **rarefaction**, with both remaining debated to date [@McMurdie2014; @Weiss2017]." - Quoted from [@Alteio2021]. Implies that rarefaction and WNWN's rarefying are the same thing.

* "**Rarefaction** is a widely used normalization technique that involves the random subsampling of sequences from the initial sample library to a selected normalized library size. This process is often dismissed as statistically invalid because subsampling effectively discards a portion of the observed sequences, yet it remains prevalent in practice and the suitability of **rarefying**, relative to many other normalization approaches, for diversity analysis has been argued" - Quoted by [@Cameron2021]. Uses rarefaction and rarefying interchangeably. This paper describes a new method as "repeated rarefying", which is the traditional rarefaction approach.

* "Confronted with technical variation as well as the overall increase in raw sequencing data generated per sample over the years, **rarefaction** (or downsampling) was suggested to standardize within and across dataset comparisons ...  However, sequencing depth-based downsizing procedures were soon criticized, not only for being wasteful and discarding information on low-abundant taxa [@McMurdie2014], but also for being unsuited when applied to communities characterized by substantial variation in cell density [@Vandeputte2017]." - Quoted by [@LlornsRico2021].  Implies that rarefaction and WNWN's rarefying are the same thing.

* "Some have suggested that data should be **rarefied** [@Weiss2017; @McKnight2018] to a common sampling depth, typically to the level of the sample with fewest sequences, while others argue that such **rarefaction** is 'inadmissible' and favor approaches that transform or scale sequence counts [@McMurdie2014; @Gloor2017]." - Quoted by [@Neu2021]. Uses rarefied and rarefaction interchangeably and implies that both are the same as WNWN's rarefying.

* "**Rarefaction** has been criticized for wasting data since we effectively remove a portion of the data in the downsampling procedure [@McMurdie2014]." - Quoted by [@Hong2022]. Here and elsewhere in this paper the authors use rarefaction and rarefying interchangeably and implies that both are the same as WNWN's rarefying.

* "**Rarefying** (also referred to as **rarefaction**) is a popular but widely criticized technique for correcting uneven sequencing depths. It involves randomly discarding counts in samples until all samples have the same predefined number of total counts." - Quoted by [@Swift2022].  As the quote indicates they use rarefying and rarefaction interchangeably although the cited WNWN indicates that they are not the same.

* "One commonly used method is to **rarefy** the data; that is, ASVs or OTUs within a sample are randomly subsampled without replacement to a preselected depth that is the same across all samples. The outcome of this is that all samples will have the same number of ASVs and any samples with fewer ASVs than the **rarefaction** level will be removed from the dataset. The level for **rarefaction** can be decided using a **rarefaction** curve, a method in which each sample is subsampled at multiple levels (e.g. 1,000 reads, 2,000 reads, 3,000 reads...), and the number of unique features or another metric of individual sample diversity of each sample at each level is measured and plotted. When the plot begins to level off after an initial climb up, the corresponding number of sequences indicates an appropriate sampling depth. The appropriate number to **rarefy** must then be balanced with the number of samples that may be dropped from the dataset which do not meet that minimum. An advantage of **rarefaction** is that it may be a more appropriate measure of very low-abundance ("rare") ASVs. This can in turn increase the accuracy of the data, as low biomass samples often have contamination and quality concerns [@Kennedy2014]. There are also disadvantages to this method, the most obvious of which is the discarding of valuable data. Clearly, this is less than ideal as the researcher must pay for the samples and sequences, and in cases where the samples are very valuable or difficult to obtain the loss of data may be destructive to the overall experimental integrity. Additionally, the loss of statistical power by removing sequences from a sample could lead to a loss of differences between two samples [@McMurdie2014]. The statistical consequences extend beyond this, as **rarefying** equalizes sample variance by adding artificial uncertainty [@McMurdie2014]." - Quoted by [@Weinroth2022]. The authors use rarefaction and rarefying interchangeably and implies that both are the same as WNWN's rarefying.

\newpage

### References

\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent

<div id="refs"></div>
\bibliography{ref}
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

\newpage